---
title: "Project 3: firstpackage Tutorial"
author: "Hannah Chang"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{firstpackage Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

This package is written for statistical inference and prediction, and includes 
the functions `my_t_test`, `my_lm`, `my_knn_cv`, and `my_rf_cv`.

To install and load this package:

```{r, eval=FALSE}
# install firstpackage
devtools::install_github("hchang23/firstpackage")

# load firstpackage
library(firstpackage)
```

# Tutorial: my_t_test

The t-tests below will be using the `lifeExp` variable from `my_gapminder` and
interpreting the results with $\alpha = 0.05$.

```{r}
# load my_gapminder data
data("my_gapminder")
```

## Test 1: alternative is "two.sided"

$H_0:\mu=60$

$H_A:\mu\neq60$

```{r}
my_t_test(my_gapminder$lifeExp, "two.sided", 60)
```

The `p_val` from the results is 0.093, which is greater than $\alpha = 0.05$,
so I fail to reject the null hypothesis. There is not enough evidence to suggest
that the mean life expectancy is not 60 years. 

## Test 2: alternative is "less"

$H_0:\mu=60$

$H_A:\mu < 60$

```{r}
my_t_test(my_gapminder$lifeExp, "less", 60)
```

The `p_val` from the results is 0.0466, which is less than $\alpha = 0.05$,
so I reject the null hypothesis. There is sufficient evidence to suggest that
the mean life expectancy is less than 60 years. 

## Test 3: alternative is "greater"

$H_0:\mu=60$

$H_A:\mu > 60$

```{r}
my_t_test(my_gapminder$lifeExp, "greater", 60)
```

The `p_val` from the results is 0.95, which is greater than $\alpha = 0.05$,
so I fail to reject the null hypothesis. There is not enough evidence to suggest
that the mean life expectancy is greater than 60 years. 

# Tutorial: my_lm



# Tutorial: my_knn_cv



# Tutorial: my_rf_cv



